---
output: html_document
---
Nonlinear models
================

Explore nonlinear models using R tools.

```{r}
require(ISLR)
attach(Wage)
```

Polynomials
------------

Focus on a single predictor, age:

```{r}
fit = lm(wage~poly(age,4),data = Wage)
summary(fit)
```

The `poly()` function generates a basis of *orthogonal polynomials*.
Let's make a plot of the fitted function, along with the standard
errors of the fit.

```{r} 
agelims = range(age)
age.grid = seq(from=agelims[1],to=agelims[2])
pred = predict(fit, newdata = list(age=age.grid), se=T)
se.bands = cbind(pred$fit+2*pred$se,pred$fit-2*pred$se)
plot(age, wage, col="darkgrey")
lines(age.grid, pred$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, col="blue", lty=2)
```

There are other ways of fitting a polynomials in R. For example:

```{r}
fita = lm(wage~age+I(age^2)+I(age^3)+I(age^4), data=Wage)
summary(fita)
```

Above `I()` wrapper is used because `age^2` means something to the formula language,
while `I(age^2)`  is protected.

The coefficients are different, but the fits are the same. 

```{r}
plot(fitted(fit), fitted(fita))
```

By using orthogonal polynomials in this simple way, it turns out 
that we can separately test for each coefficients. So we look at the summary again,
we can see that the linear, quadratic, and cubic terms are significant, but not the quartic.

This only works  with linear regression, and if there is a single predictor.
In general, we would use `anova()` as this next example demonstrates.
This is an example of nested sequence of complexity:

```{r}
fita = lm(wage~education, data=Wage)
fitb = lm(wage~education+age, data=Wage)
fitc = lm(wage~education+poly(age,2), data=Wage)
fitd = lm(wage~education+poly(age,3), data=Wage)
anova(fita, fitb, fitc, fitd)
```

### Polynomial logistic regression

Now we fit a logistic regression model to a binary response variable,
constructed from `wage`. We code the big earners (`>250k`) as 1, else 0

```{r}
fita = glm(I(wage > 250) ~ poly(age,2), data=Wage, family=binomial)
fitb = glm(I(wage > 250) ~ poly(age,3), data=Wage, family=binomial)
anova(fita, fitb)
summary(fit)
preds = predict(fitb, list(age=age.grid), se=T)
se.bands = preds$fit + cbind(fit=0, lower=-2*preds$se, upper=2*preds$se)
se.bands[1:5,]
```

The above calculations are on the logit scale. We need to transform it back using
the inverse logit $$p=\frac{e^\eta}{1+e^\eta}.$$ We can do this simultaneously for all 
three columns of `se.bands`:

```{r}
prob.bands = exp(se.bands)/(1+exp(se.bands))
matplot(age.grid, prob.bands, col="blue", lwd=c(2,1,1), lty=c(1,2,2), type="l", ylim=c(0,0.1))
points(jitter(age), I(wage>250)/10, pch="|", cex=0.5)
```




