---
output: html_document
---
Nonlinear models: Polynomials and Splines
=========================================

Explore nonlinear models using R tools.

```{r}
require(ISLR)
attach(Wage)
```

Polynomials
------------

Focus on a single predictor, age:

```{r}
fit = lm(wage~poly(age,4),data = Wage)
summary(fit)
```

The `poly()` function generates a basis of *orthogonal polynomials*.
Let's make a plot of the fitted function, along with the standard
errors of the fit.

```{r} 
agelims = range(age)
age.grid = seq(from=agelims[1],to=agelims[2])
pred = predict(fit, newdata = list(age=age.grid), se=T)
se.bands = cbind(pred$fit+2*pred$se,pred$fit-2*pred$se)
plot(age, wage, col="darkgrey")
lines(age.grid, pred$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, col="blue", lty=2)
```

There are other ways of fitting a polynomials in R. For example:

```{r}
fita = lm(wage~age+I(age^2)+I(age^3)+I(age^4), data=Wage)
summary(fita)
```

Above `I()` wrapper is used because `age^2` means something to the formula language,
while `I(age^2)`  is protected.

The coefficients are different, but the fits are the same. 

```{r}
plot(fitted(fit), fitted(fita))
```

By using orthogonal polynomials in this simple way, it turns out 
that we can separately test for each coefficients. So we look at the summary again,
we can see that the linear, quadratic, and cubic terms are significant, but not the quartic.

This only works  with linear regression, and if there is a single predictor.
In general, we would use `anova()` as this next example demonstrates.
This is an example of nested sequence of complexity:

```{r}
fita = lm(wage~education, data=Wage)
fitb = lm(wage~education+age, data=Wage)
fitc = lm(wage~education+poly(age,2), data=Wage)
fitd = lm(wage~education+poly(age,3), data=Wage)
anova(fita, fitb, fitc, fitd)
```

### Polynomial logistic regression

Now we fit a logistic regression model to a binary response variable,
constructed from `wage`. We code the big earners (`>250k`) as 1, else 0

```{r}
fita = glm(I(wage > 250) ~ poly(age,2), data=Wage, family=binomial)
fitb = glm(I(wage > 250) ~ poly(age,3), data=Wage, family=binomial)
anova(fita, fitb)
summary(fit)
preds = predict(fitb, list(age=age.grid), se=T)
se.bands = preds$fit + cbind(fit=0, lower=-2*preds$se, upper=2*preds$se)
se.bands[1:5,]
```

The above calculations are on the logit scale. We need to transform it back using
the inverse logit $$p=\frac{e^\eta}{1+e^\eta}.$$ We can do this simultaneously for all 
three columns of `se.bands`:

```{r}
prob.bands = exp(se.bands)/(1+exp(se.bands))
matplot(age.grid, prob.bands, col="blue", lwd=c(2,1,1), lty=c(1,2,2), type="l", ylim=c(0,0.1))
points(jitter(age), I(wage>250)/10, pch="|", cex=0.5)
```

Splines
-------

Splines are more flexible than polynomials but the idea is rather similar. Here we will explore 
cubic splines.

```{r}
require(splines)
fit = lm(wage~bs(age, knots=c(25,40,60)), data=Wage)
summary(fit)
plot(age, wage, col="darkgrey")
lines(age.grid, predict(fit, list(age=age.grid)), col="darkgreen", lwd=2)
abline(v=c(25,40,60), lty=2, col="darkgreen")
```


The smoothing splines don't require knot selection, but they do have a smoothing parameter,
which can be conveniently specified via the effective degrees of freedom or `df`.

```{r}
fit = smooth.spline(age, wage, df=16)
plot(age, wage, col="darkgrey")
lines(fit, col="red", lwd=2)
```

Or we can use LOO cross-validation  to select the smoothing parameter for us
automatically:

```{r}
fit = smooth.spline(age, wage, cv=T)
plot(age, wage, col="darkgrey")
lines(fit, col="purple", lwd=2)
fit
```

Generalized Additive Models
---------------------------
So far we have focused on fitting models with mostly a single nonlinear term.
The `gam` package makes it easy to work with multiple nonlinear terms.
In addition it knows how to plot these functions and their standard errors.


```{r fig.width=10, fig.height=4}
require(gam)
gam1 = gam(wage~s(age,df=4)+s(year,df=4)+education,data = Wage)
par(mfrow=c(1,3))
plot(gam1,se=T)
```

Let's do the same, but now with logistic regression:

```{r fig.width=10, fig.height=4}
require(gam)
gam2 = gam(I(wage>250)~s(age,4)+s(year,4)+education, data=Wage, family = binomial)
par(mfrow=c(1,3))
plot(gam2)
```


Let's see if we need a nonlinear terms for year

```{r}
gam2a = gam(I(wage>250)~s(age,4)+year+education, data=Wage, family=binomial)
anova(gam2a,gam2,test="Chisq")
```

The p-value is very high the nonlinear `year` term, so we don't need it. We may not even need `year`.

Nice feature of `gam` is that it knows how to plot the functions nicely, 
even for models fit by `lm` and `glm`.

```{r fig.width=10, fig.height=4}
par(mfrow=c(1,3))
lm1 = lm(wage~ns(age,df=4)+ns(year,df=4)+education,data=Wage)
plot.gam(lm1, se=T)
```

See the chapter on `gam` for more in Introduction to Statistical Learning.
