---
title: "Trees-Random-Forest-Boosting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Decision Tree
=============

We will examine the `Carseats` data using the `tree` package in R, 
as in the lab in the book (Introduction to Statistical Learning with R).

We create a binary response variable `High` (for high sales) and we include
it in the same dataframe.

```{r}
require(ISLR)
require(tree)
attach(Carseats)
View(Carseats)
hist(Sales)
High = ifelse(Sales<=8, "No", "Yes")
Carseats = data.frame(Carseats, High)
```

Now, we fit a tree to these data, and summarize and plot it. Notice we have to _exclude_ `Sales` from the right-hand side of the formula, because the response is derived from it. This fit is the simplest possible,
there are two possible outcomes.

```{r}
tree.carseats = tree(High~.-Sales, data=Carseats)
summary(tree.carseats)
plot(tree.carseats)
text(tree.carseats, pretty=0)
```

For a detailed summary of the tree, print it:

```{r}
tree.carseats
```

Let's create training and test sets (250, 150) of the 400 observations.


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
